---
title: "Maxent_Input_Wrangling"
output: html_document
---

**This script cleans all environmental layers for the three time periods (2005-2015, 2040-2050, 2070-2080) for use in Maxent through cropping, resampling, and exporting as ASCII files**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load_packages, message=FALSE}

library(raster)
library(tidyverse)
library(dplyr)
library(here)
library(ascii)
library(maptools)
#library(sf)
library(fasterize)
library(rgdal)
library(ncdf4)
library(rgeos)
library(maptools)
library(tmap)

```

###read in data
```{r}

#PRESENT
pH_present <- raster ("G:/Coral_Suitability/Maxent/Maxent_Inputs/Present_ObservationData/pH_GMEDcurrent.tif")
par_present <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/Present_ObservationData/PAR_Present_Bio.tif")
salinity_present <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/Present_ObservationData/Salinity_Present_Bio.tif")
SST_present <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/Present_ObservationData/SST_Present_Bio.tif")
currentspeed_present <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/Present_ObservationData/UVEL_Present_Bio.tif")

##### RCP 4.5 #####

#NEARFUTURE4.5 (2040-2050)
pH_near45 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP4_5/2050/ph_near45.asc")
par_near45 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP4_5/2050/par_near45.asc")
salinity_near45 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP4_5/2050/salinity_near45.asc")
SST_near45 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP4_5/2050/sst_near45.asc")
currentspeed_near45 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP4_5/2050/speed_near45.asc")


#FUTURE4.5 (2070-2080)
pH_future45 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP4_5/2080/ph_far45.asc")
par_future45 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP4_5/2080/par_far45.asc")
salinity_future45 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP4_5/2080/salinity_far45.asc")
SST_future45 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP4_5/2080/sst_far45.asc")
currentspeed_future45 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP4_5/2080/speed_far45.asc")

##### RCP 8.5 #####

#NEARFUTURE8.5 (2040-2050)
pH_near85 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP8_5/2050/ph_nearfuture.asc")
par_near85 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP8_5/2050/par_nearfuture.asc")
salinity_near85 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP8_5/2050/salinity_nearfuture.asc")
SST_near85 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP8_5/2050/sst_nearfuture.asc")
currentspeed_near85 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP8_5/2050/speed_nearfuture.asc")


#FUTURE8.5 (2070-2080)
pH_future85 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP8_5/2080/ph_future85.asc")
par_future85 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP8_5/2080/par_future85.asc")
salinity_future85 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP8_5/2080/salinity_future85.asc")
SST_future85 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP8_5/2080/sst_future85.asc")
currentspeed_future85 <- raster("G:/Coral_Suitability/Maxent/Maxent_Inputs/ASCII_Downscaled_Files/RCP8_5/2080/speed_future85.asc")


```

###set Caribbean extent
```{r set_extent}

Caribbean_extent <- as(extent(-103, -50, 4, 30), 'SpatialPolygons')

```


###crop present, nearfuture, and future pH rasters
```{r}

#PRESENT
pH_present <- crop(pH_present, Caribbean_extent)
plot(pH_present)

##### RCP 4.5 #####

#NEAR FUTURE
pH_near45 <- crop(pH_near45, Caribbean_extent)
plot(pH_near45)

#FUTURE
pH_future45 <- crop(pH_future45, Caribbean_extent)
plot(pH_future45)

##### RCP 8.5 #####


#NEAR FUTURE
pH_near85 <- crop(pH_near85, Caribbean_extent)
plot(pH_near85)

#FUTURE
pH_future85 <- crop(pH_future85, Caribbean_extent)
plot(pH_future85)

summary(pH_present)
summary(pH_near45)
summary(pH_future45)
summary(pH_near85)
summary(pH_future85)

```

###crop present, nearfuture, and future SST rasters
```{r}

#PRESENT
SST_present <- crop(SST_present, Caribbean_extent)
plot(SST_present)

##### RCP 4.5 #####

#NEAR FUTURE
SST_near45 <- crop(SST_near45, Caribbean_extent)
plot(SST_near45)

#FUTURE
SST_future45 <- crop(SST_future45, Caribbean_extent)
plot(SST_future45)

##### RCP 8.5 #####

#NEAR FUTURE
SST_near85 <- crop(SST_near85, Caribbean_extent)
plot(SST_near85)

#FUTURE
SST_future85 <- crop(SST_future85, Caribbean_extent)
plot(SST_future85)

summary(SST_present)
summary(SST_near45)
summary(SST_future45)
summary(SST_near85)
summary(SST_future85)

```

###crop present, nearfuture, and future PAR rasters
```{r}

#PRESENT
par_present <- crop(par_present, Caribbean_extent)
plot(par_present)

##### RCP 4.5 #####

#NEAR FUTURE
par_near45 <- crop(par_near45, Caribbean_extent)
plot(par_near45)

#FUTURE
par_future45 <- crop(par_future45, Caribbean_extent)
plot(par_future45)

##### RCP 8.5 #####

#NEAR FUTURE
par_near85 <- crop(par_near85, Caribbean_extent)
plot(par_near85)

#FUTURE
par_future85 <- crop(par_future85, Caribbean_extent)
plot(par_future85)

summary(par_present)
summary(par_near45)
summary(par_future45)
summary(par_near85)
summary(par_future85)

```

###crop present, nearfuture, and future salinity rasters
```{r}

#PRESENT
salinity_present <- crop(salinity_present, Caribbean_extent)
plot(salinity_present)

##### RCP 4.5 #####

#NEAR FUTURE
salinity_near45 <- crop(salinity_near45, Caribbean_extent)
plot(salinity_near45)

#FUTURE
salinity_future45 <- crop(salinity_future45, Caribbean_extent)
plot(salinity_future45)

##### RCP 8.5 #####

#NEAR FUTURE
salinity_near85 <- crop(salinity_near85, Caribbean_extent)
plot(salinity_near85)

#FUTURE
salinity_future85 <- crop(salinity_future85, Caribbean_extent)
plot(salinity_future85)

summary(salinity_present)
summary(salinity_near45)
summary(salinity_future45)
summary(salinity_near85)
summary(salinity_future85)

```

###crop present, nearfuture, and future current speed rasters
```{r}

#PRESENT
currentspeed_present <- crop(currentspeed_present, Caribbean_extent)
plot(currentspeed_present)

##### RCP 4.5 #####

#NEAR FUTURE
currentspeed_near45 <- crop(currentspeed_near45, Caribbean_extent)
plot(currentspeed_near45)

#FUTURE
currentspeed_future45 <- crop(currentspeed_future45, Caribbean_extent)
plot(currentspeed_future45)

##### RCP 8.5 #####

currentspeed_near85 <- crop(currentspeed_near85, Caribbean_extent)
plot(currentspeed_near85)

#FUTURE
currentspeed_future85 <- crop(currentspeed_future85, Caribbean_extent)
plot(currentspeed_future85)

summary(currentspeed_present)
summary(currentspeed_near45)
summary(currentspeed_near85)
summary(currentspeed_future45)
summary(currentspeed_future85)


```

###resample Caribbean data for Maxent
```{r resample_files}

#PRESENT
SST_present_resample <- resample(SST_present, pH_present)
pH_present_resample <- resample(pH_present, pH_present)
salinity_present_resample <- resample(salinity_present, pH_present)
currentspeed_present_resample <- resample(currentspeed_present, pH_present)
par_present_resample <- resample(par_present, pH_present)

##### RCP 4.5 #####

#NEAR FUTURE
SST_near45_resample <- resample(SST_near45, pH_present)
salinity_near45_resample <- resample(salinity_near45, pH_present)
currentspeed_near45_resample <- resample(currentspeed_near45, pH_present)
par_near45_resample <- resample(par_near45, pH_present)
pH_near45_resample <- resample(pH_near45, pH_present)

#FUTURE
SST_future45_resample <- resample(SST_future45, pH_present)
salinity_future45_resample <- resample(salinity_future45, pH_present)
currentspeed_future45_resample <- resample(currentspeed_future45, pH_present)
par_future45_resample <- resample(par_future45, pH_present)
pH_future45_resample <- resample(pH_future45, pH_present)

##### RCP 8.5 #####

#NEAR FUTURE
SST_near85_resample <- resample(SST_near85, pH_present)
salinity_near85_resample <- resample(salinity_near85, pH_present)
currentspeed_near85_resample <- resample(currentspeed_near85, pH_present)
par_near85_resample <- resample(par_near85, pH_present)
pH_near85_resample <- resample(pH_near85, pH_present)

#FUTURE
SST_future85_resample <- resample(SST_future85, pH_present)
salinity_future85_resample <- resample(salinity_future85, pH_present)
currentspeed_future85_resample <- resample(currentspeed_future85, pH_present)
par_future85_resample <- resample(par_future85, pH_present)
pH_future85_resample <- resample(pH_future85, pH_present)

```

###export data as ASCII files for Maxent
```{r export_data}

#PRESENT
writeRaster(pH_present_resample, filename = "pH_present.asc")
writeRaster(SST_present_resample, filename = "SST_present.asc")
writeRaster(salinity_present_resample, filename = "salinity_present.asc")
writeRaster(currentspeed_present_resample, filename = "currentspeed_present.asc")
writeRaster(par_present_resample, filename = "PAR_present.asc")

##### RCP 4.5 #####

#NEAR FUTURE
writeRaster(SST_near45_resample, filename = "sst_near45.asc")
writeRaster(salinity_near45_resample, filename = "salinity_near45.asc")
writeRaster(currentspeed_near45_resample, filename = "currentspeed_near45.asc")
writeRaster(par_near45_resample, filename = "PAR_near45.asc")
writeRaster(pH_near45_resample, filename = "pH_near45.asc")

#FUTURE
writeRaster(SST_future45_resample, filename = "SST_future45.asc")
writeRaster(salinity_future45_resample, filename = "salinity_future45.asc")
writeRaster(currentspeed_future45_resample, filename = "currentspeed_future45.asc")
writeRaster(par_future45_resample, filename = "PAR_future45.asc")
writeRaster(pH_future45_resample, filename = "pH_future45.asc")

##### RCP 8.5 #####

#NEAR FUTURE
writeRaster(SST_near85_resample, filename = "sst_near85.asc")
writeRaster(salinity_near85_resample, filename = "salinity_near85.asc")
writeRaster(currentspeed_near85_resample, filename = "currentspeed_near85.asc")
writeRaster(par_near85_resample, filename = "PAR_near85.asc")
writeRaster(pH_near85_resample, filename = "pH_near85.asc")

#FUTURE
writeRaster(SST_future85_resample, filename = "SST_future85.asc")
writeRaster(salinity_future85_resample, filename = "salinity_future85.asc")
writeRaster(currentspeed_future85_resample, filename = "currentspeed_future85.asc")
writeRaster(par_future85_resample, filename = "PAR_future85.asc")
writeRaster(pH_future85_resample, filename = "pH_future85.asc")

```



